Investigating soundscapes perception through acoustic scenes simulation.

### 2018-01-29 lafaySimulatedScenePerception.tex
Réponses aux différents commentaires

Accounting for the previously detailed perceptual matters, two taxonomies are built, one for event sounds and one for texture sounds. Four \ml{Three ?} levels of abstraction are considered from the most generic (level 0) to the most specific classes (level 2), leading to a taxonomy close to the one used in \cite{Salamon14}. Only three levels of abstraction are considered for the texture sounds.

@@

All samples are normalized to the same $RMS$ level of $-12$ $dB$ FS, \ie relative to Full Scale. In our case, the full scale level is set arbitrarily to 1 Volt.

@@

\jfp{je ne comprends pas, pour moi il faut faire, pour chaque sujet de l'expé 1, un test de différence entre Aniscène  et Aiscène
a quoi sert Asubject ici ?}

@@

\jfp{a la limite on pourrait ne pas écarter arbitrairement de scène et ne garder que celles pour lesquelles les iscenes sont signif. + plaisante que les ni-scene. ca revient au même. c'est les test qui les écarte. il faudrait peut être justifier pourquoi on fait cela  : pour ne pas fausser les résultats de l'analyse et vérifier que les i scènes sont effectivement perçues + plaisantes que les ni.}

@@

It appears that the biggest influence on the global sound levels comes from the events, the difference between $L$ and $L(E)$ being only 1 $dB$ for i- and ni-scenes. This observation is in agreement with the results obtained by Kuwano~\al~\cite{kuwano_memory_2003}. During their experiment, the authors ask their subjects to assess a set of soundscapes at a global level and then to do the same judgment at the time when they detect a sound source. The study shows that there is no significant difference between global and averaged instantaneous judgments. In our case, the result can be interpreted as if the subjects had unconsciously integrated this perceptual reality when composing the scenes, by allocating most of the global sound levels to well identified and relatively short sounds, \ie the events.

@@

  \subfloat[]{\includegraphics[width=.5\linewidth]{gfx/xp1_class_2}\label{fig:soundsourceb}}
  \subfloat[]{\includegraphics[width=.5\linewidth]{gfx/xp1_class_3}\label{fig:soundsourcec}}
                             \caption{Proportion of simulated scenes (i or ni) involving a specific class of sounds: (a) event classes at an abstraction level $0+$, (b) texture classes at an abstraction level $0$, (c) event sub-classes at an abstraction level $1$ belonging to \emph{traffic} and \emph{public transportation} classes of the abstraction level $0$. \jfp{pourquoi la classe "start passage way", présente en figure A9, n'est pas présente sur ce graphe ?}}\label{fig:soundsource}
\end{figure}

@@

Considering $E$ and $T$ separately, it appears that: 1) the $p@5$ obtained with $E$ is similar to the one obtained with $ET$ ; 2) the $p@5$ obtained with $T$ is always lower than the one obtained with $E$, by $10\%$ to $15\%$. Those results indicate that the semantic information that is discriminative is mostly carried by the events. Those results are in line with \cite{maffiolo_caracterisation_1999}. As discussed in Section~\ref{sec:simscene_sampleDataSet}, it seems that humans analyze the event scenes which are composed of several sound events in a causal manner, \ie by identifying the sources.

@@

\ml{Ma traduction : Another hypothesis would be that the volume somehow act as an hedonic "gain" factor. If the volume of a negative marker is high, it lowers the overall pleasantness. If the volume of a positive marker is high, it highers the overall pleasantness. Evidently, the positive gain is expected to saturate at a given level and will quickly decreases as the level raises above a given threshold.}

@@

The aim here is to validate the hypothesis that the removal of i-markers and ni-markers have an impact on the perceived pleasantness. To do so, we perform a variance analysis. We consider $\mathcal{A}_{subject}$ as a dependent variable, and as independent variables, the type of environment (i/ni) and the presence/absence of markers (km/rm). As each subject evaluated the whole set of scenes, a two factors repeated measure ANOVA is used to evaluate whether there exist a significant difference of perceived pleasantness between km-scenes and rm-scenes. The two independent variables are considered as \emph{within-subject} factors. The factors being of only two levels each (type: i/ni,  marker: km/rm), the sphericity hypothesis does not need to be checked. \emph{Post hoc} analyses are done using the Tukey-Kramer procedure. \jfp{l'analyse est inutile a mon avis, il n'y a que 2 niveaux pour chaque facteur de l'ANOVA.}

>>>>

Accounting for the previously detailed perceptual matters, two taxonomies are built, one for event sounds and one for texture sounds. Four \ml{Three ?} \gl{non 4} levels of abstraction are considered from the most generic (level 0) to the most specific classes (level 2), leading to a taxonomy close to the one used in \cite{Salamon14}. Only three levels of abstraction are considered for the texture sounds.

@@

All samples are normalized to the same $RMS$ level of $-12$ $dB$ FS, \ie~relative to Full Scale. In our case, the full scale level is set arbitrarily to 1 Volt.

@@

\jfp{je ne comprends pas, pour moi il faut faire, pour chaque sujet de l'expé 1, un test de différence entre Aniscène  et Aiscène a quoi sert Asubject ici ?} \gl{Les deux sont possibles : Ascene mesure des différences structurelles, Asubject des différences perceptives}

@@

\jfp{a la limite on pourrait ne pas écarter arbitrairement de scène et ne garder que celles pour lesquelles les iscenes sont signif. + plaisante que les ni-scene. ca revient au même. c'est les test qui les écarte. il faudrait peut être justifier pourquoi on fait cela  : pour ne pas fausser les résultats de l'analyse et vérifier que les i scènes sont effectivement perçues + plaisantes que les ni.} \gl{Oui. Mais ça revient au même de toute façon}

@@

It appears that the biggest influence on the global sound levels comes from the events, the difference between $L$ and $L(E)$ being only 1 $dB$ for i- and ni-scenes. This observation is in agreement with the results obtained by Kuwano~\al~\cite{kuwano_memory_2003}. During their experiment, the authors ask their subjects to assess a set of soundscapes at a global level and then to do the same judgment at the time when they detect a sound source. The study shows that there is no significant difference between global and averaged instantaneous judgments. In our case, the result can be interpreted as if the subjects had unconsciously integrated this perceptual reality when composing the scenes, by allocating most of the global sound levels to well identified and relatively short sounds, \ie~the events.

@@

  \subfloat[]{\includegraphics[width=.5\linewidth]{gfx/xp1_class_2}\label{fig:soundsourceb}}
  \subfloat[]{\includegraphics[width=.5\linewidth]{gfx/xp1_class_3}\label{fig:soundsourcec}}
                             \caption{Proportion of simulated scenes (i or ni) involving a specific class of sounds: (a) event classes at an abstraction level $0+$, (b) texture classes at an abstraction level $0$, (c) event sub-classes at an abstraction level $1$ belonging to \emph{traffic} and \emph{public transportation} classes of the abstraction level $0$. \jfp{pourquoi la classe "start passage way", présente en figure A9, n'est pas présente sur ce graphe ?} \gl{Elle n'est pas parlante}}\label{fig:soundsource}
\end{figure}

@@

Considering $E$ and $T$ separately, it appears that: 1) the $p@5$ obtained with $E$ is similar to the one obtained with $ET$ ; 2) the $p@5$ obtained with $T$ is always lower than the one obtained with $E$, by $10\%$ to $15\%$. Those results indicate that the semantic information that is discriminative is mostly carried by the events. Those results are in line with \cite{maffiolo_caracterisation_1999}. As discussed in Section~\ref{sec:simscene_sampleDataSet}, it seems that humans analyze the event scenes which are composed of several sound events in a causal manner, \ie~by identifying the sources.

@@

\ml{Ma traduction : Another hypothesis would be that the volume somehow act as an hedonic "gain" factor. If the volume of a negative marker is high, it lowers the overall pleasantness. If the volume of a positive marker is high, it highers the overall pleasantness. Evidently, the positive gain is expected to saturate at a given level and will quickly decreases as the level raises above a given threshold.} \gl{@mathieu, pourquoi pas. Mais je pense qu'on peut laisser également le premier paragraphe, ça permettra aux reviewer de commenter qqch :). Quid de mettre les deux.}

@@

The aim here is to validate the hypothesis that the removal of i-markers and ni-markers have an impact on the perceived pleasantness. To do so, we perform a variance analysis. We consider $\mathcal{A}_{subject}$ as a dependent variable, and as independent variables, the type of environment (i/ni) and the presence/absence of markers (km/rm). As each subject evaluated the whole set of scenes, a two factors repeated measure ANOVA is used to evaluate whether there exist a significant difference of perceived pleasantness between km-scenes and rm-scenes. The two independent variables are considered as \emph{within-subject} factors. The factors being of only two levels each (type: i/ni,  marker: km/rm), the sphericity hypothesis does not need to be checked. \emph{Post hoc} analyses are done using the Tukey-Kramer procedure. \jfp{l'analyse est inutile a mon avis, il n'y a que 2 niveaux pour chaque facteur de l'ANOVA.} \gl{2 niveaux par facteur, mais 4 niveaux en tout}
